---
title: "Core data processing and analysis pipeline"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=F, message=F, warning = F)
```

## Load packages

```{r eval=T}
library(tidyverse)
library(cowplot)
```

## Come up with some sample lists and tables

```{r eval=T}
base_dir = "/workdir/batch-effect/"
# Read in the full sample table from the Greenland cod project
sample_table_full <- read_tsv("../../cod/greenland-cod/sample_lists/sample_table_merged_mincov_contamination_filtered.tsv") %>%
  bind_cols(bam_list = read_lines("../../cod/greenland-cod/sample_lists/bam_list_realigned_mincov_contamination_filtered.txt"))
# Select a subset of populations that were sequenced in both Hiseq and Nextseq platforms
# Note that there are some QQL samples that appear to be "merged", but they were merged from lane 1 and 2
sample_table_merged <- filter(sample_table_full, 
                                    population %in% c("IKE2011", "QQL2011", "ITV2011", "KNG2011", "BUK2011", "NAR2008", "UUM2010", "PAA2011", "ATP2011")) %>%
  filter(data_type != "pese")
## Raw bam file list for the pair end samples
bam_list_merged_pe <- sample_table_merged %>%
  filter(data_type=="pe") %>%
  mutate(bam_list = str_c(base_dir, "bam/", sample_seq_id, "_bt2_gadMor3_sorted.bam")) %>%
  dplyr::select(bam_list)
## Raw bam file list for the single end samples
bam_list_merged_se <- sample_table_merged %>%
  filter(data_type=="se") %>%
  mutate(bam_list = str_c(base_dir, "bam/", sample_seq_id, "_bt2_gadMor3_sorted.bam")) %>%
  dplyr::select(bam_list)
## Overlap clipped bamlist for all samples
bam_list_overlap_clipped <- sample_table_merged %>%
  mutate(bam_list = ifelse(data_type == "pe", 
                           str_c(base_dir, "bam/", sample_seq_id, "_bt2_gadMor3_sorted_dedup_overlapclipped.bam"),
                           str_c(base_dir, "bam/", sample_seq_id, "_bt2_gadMor3_sorted_dedup.bam"))) %>%
  dplyr::select(bam_list)
## Indel realigned bamlist for all samples
bam_list_realigned <- sample_table_merged %>%
  mutate(bam_list = ifelse(data_type == "pe", 
                           str_c(base_dir, "bam/", sample_seq_id, "_bt2_gadMor3_sorted_dedup_overlapclipped_realigned.bam"),
                           str_c(base_dir, "bam/", sample_seq_id, "_bt2_gadMor3_sorted_dedup_realigned.bam"))) %>%
  dplyr::select(bam_list, data_type)

## Get unmerged sample table for the select subset of samples
sample_table_unmerged <- read_tsv("../../cod/greenland-cod/sample_lists/sample_table.tsv") %>%
  semi_join(sample_table_merged, by=c("sample_id"="sample_id_corrected"))
## Fastq list of pe and se samples
fastq_list_pe <- filter(sample_table_unmerged, lane_number == 7)$prefix
fastq_list_se <- filter(sample_table_unmerged, lane_number != 7)$prefix

## Sample distribution
sample_size_plot <- sample_table_merged %>%
  mutate(sample_id_corrected=fct_reorder(sample_id_corrected, data_type)) %>%
  ggplot(aes(x=population_new, fill=data_type, group=sample_id_corrected)) +
  geom_bar(color="black") +
  xlab("population")+
  ylab("sample size") +
  theme_cowplot() +
  coord_flip() +
  theme(legend.position = "none")
sample_size_plot
## Number of bases
base_count <- read_tsv("../sample_lists/count_merged_old.tsv") %>%
  dplyr::select(sample_id_corrected, final_mapped_bases)
coverage_plot <- sample_table_merged %>%
  left_join(base_count) %>%
  arrange(data_type, desc(final_mapped_bases)) %>%
  mutate(sample_id_corrected=as_factor(sample_id_corrected)) %>%
  ggplot(aes(x=population_new, y=final_mapped_bases/0.67/10^9, fill=data_type, group=sample_id_corrected)) +
  geom_col(color="black") +
  scale_fill_discrete(labels=c("NextSeq-150PE", "HiSeq-125SE")) +
  labs(x="population", y="coverage", fill="batch")+
  ylab("coverage") +
  theme_cowplot() +
  coord_flip()
coverage_plot
sample_table_merged %>%
  left_join(base_count) %>%
  ggplot(aes(x=final_mapped_bases, fill=data_type)) +
  geom_histogram(position="dodge") +
  theme_cowplot()
sample_table_merged %>%
  left_join(base_count) %>%
  group_by(data_type) %>%
  summarise(sample_size=n(), avearge_final_coverage=mean(final_mapped_bases)/0.67/10^9)
```

```{r eval=TRUE, fig.width=15, fig.height=4}
cowplot::plot_grid(sample_size_plot, coverage_plot, nrow = 1, rel_widths = c(1, 1.2))
```

```{r eval=F}
# Write the objects created above
sample_table_merged %>%
  dplyr::select(-bam_list) %>%
  write_tsv("../sample_lists/sample_table_merged.tsv")
write_tsv(bam_list_merged_pe, "../sample_lists/bam_list_merged_pe.txt", col_names = F)
write_tsv(bam_list_merged_se, "../sample_lists/bam_list_merged_se.txt", col_names = F)
write_tsv(bam_list_overlap_clipped, "../sample_lists/bam_list_dedup_overlapclipped.txt", col_names = F)
bam_list_realigned %>%
  dplyr::select(bam_list) %>%
  write_tsv("../sample_lists/bam_list_realigned.txt", col_names = F)
bam_list_realigned %>%
  filter(data_type=="pe") %>%
  dplyr::select(bam_list) %>%
  write_tsv("../sample_lists/bam_list_per_pop/bam_list_realigned_pe.txt", col_names = F)
bam_list_realigned %>%
  filter(data_type=="se") %>%
  dplyr::select(bam_list) %>%
  write_tsv("../sample_lists/bam_list_per_pop/bam_list_realigned_se.txt", col_names = F)
write_tsv(sample_table_unmerged, "../sample_lists/sample_table_unmerged.tsv")
write_lines(fastq_list_pe, "../sample_lists/fastq_list_pe.txt")
write_lines(fastq_list_se, "../sample_lists/fastq_list_se.txt")
```

## Process the PE data

#### polyG trimming (`--trim_poly_g`)

```{bash eval=F}
echo 'BASEDIR=/workdir/batch-effect/
INPUTDIR=/workdir/cod/greenland-cod/adapter_clipped/
SAMPLELIST=$BASEDIR/sample_lists/fastq_list_pe.txt
SAMPLETABLE=$BASEDIR/sample_lists/sample_table_unmerged.tsv

for SAMPLEFILE in `cat $SAMPLELIST`; do
  SAMPLE_ID=`grep -P "${SAMPLEFILE}\t" $SAMPLETABLE | cut -f 4`
  SEQ_ID=`grep -P "${SAMPLEFILE}\t" $SAMPLETABLE | cut -f 3`
  LANE_ID=`grep -P "${SAMPLEFILE}\t" $SAMPLETABLE | cut -f 2`
  SAMPLE_SEQ_ID=$SAMPLE_ID"_"$SEQ_ID"_"$LANE_ID
  
  ## The input and output path and file prefix
  SAMPLEADAPT=$INPUTDIR$SAMPLE_SEQ_ID
  SAMPLEQUAL=$BASEDIR"polyg_trimmed/"$SAMPLE_SEQ_ID

  /workdir/programs/fastp --trim_poly_g -L -A \
  -i $SAMPLEADAPT"_adapter_clipped_f_paired.fastq.gz" \
  -I $SAMPLEADAPT"_adapter_clipped_r_paired.fastq.gz" \
  -o $SAMPLEQUAL"_adapter_clipped_qual_filtered_f_paired.fastq.gz" \
  -O $SAMPLEQUAL"_adapter_clipped_qual_filtered_r_paired.fastq.gz" \
  -h $SAMPLEQUAL"_adapter_clipped_qual_filtered_fastp.html"
done' > /workdir/batch-effect/scripts/trim_polyg_batch_effect.sh

nohup bash /workdir/batch-effect/scripts/trim_polyg_batch_effect.sh > /workdir/batch-effect/nohups/cut_right_batch_effect.nohups &
```

#### Sliding window trimming (`--cut_right`)

```{bash eval=F}
echo 'SAMPLELIST=/workdir/cod/greenland-cod/sample_lists/sample_list_pe_1.tsv 
SAMPLETABLE=/workdir/cod/greenland-cod/sample_lists/sample_table_pe.tsv
BASEDIR=/workdir/batch-effect/
INPUTDIR=/workdir/cod/greenland-cod/adapter_clipped/

for SAMPLEFILE in `cat $SAMPLELIST`; do
  SAMPLE_ID=`grep -P "${SAMPLEFILE}\t" $SAMPLETABLE | cut -f 4`
  SEQ_ID=`grep -P "${SAMPLEFILE}\t" $SAMPLETABLE | cut -f 3`
  LANE_ID=`grep -P "${SAMPLEFILE}\t" $SAMPLETABLE | cut -f 2`
  SAMPLE_SEQ_ID=$SAMPLE_ID"_"$SEQ_ID"_"$LANE_ID
  
  ## The input and output path and file prefix
  SAMPLEADAPT=$INPUTDIR$SAMPLE_SEQ_ID
  SAMPLEQUAL=$BASEDIR"cut_right/"$SAMPLE_SEQ_ID

  /workdir/programs/fastp --trim_poly_g -L -A --cut_right \
  -i $SAMPLEADAPT"_adapter_clipped_f_paired.fastq.gz" \
  -I $SAMPLEADAPT"_adapter_clipped_r_paired.fastq.gz" \
  -o $SAMPLEQUAL"_adapter_clipped_qual_filtered_f_paired.fastq.gz" \
  -O $SAMPLEQUAL"_adapter_clipped_qual_filtered_r_paired.fastq.gz" \
  -h $SAMPLEQUAL"_adapter_clipped_qual_filtered_fastp.html"
done' > /workdir/batch-effect/scripts/cut_right_batch_effect.sh

nohup bash /workdir/batch-effect/scripts/cut_right_batch_effect.sh > /workdir/batch-effect/nohups/cut_right_batch_effect.nohups &
```

#### Map to reference genome

This step also includes quality filtering and sorting, but I will rerun the quality filtering and sorting step with no quality filter, both for PE and SE data. 

```{bash eval=F}
nohup bash /workdir/data-processing/scripts/low_coverage_mapping.sh \
/workdir/batch-effect/sample_lists/fastq_list_pe.txt \
/workdir/batch-effect/sample_lists/sample_table_unmerged.tsv \
/workdir/batch-effect/cut_right/ \
/workdir/batch-effect/ \
_adapter_clipped_qual_filtered_f_paired.fastq.gz \
_adapter_clipped_qual_filtered_r_paired.fastq.gz \
very-sensitive \
/workdir/cod/reference_seqs/gadMor3.fasta \
gadMor3 \
> /workdir/batch-effect/nohups/low_coverage_mapping_pe.nohup &
```

## Process the PE and SE data together

These steps are necessary for the SE data because intermediate bam files were deleted. 

#### Sort the raw bam files

Save the following script as `/workdir/batch-effect/scripts/sort_raw_bam.sh`

```{bash eval=F}
OUTPUTPATH='/workdir/batch-effect/bam/'

for K in {2..168}; do
  SAMPLE=`head /workdir/batch-effect/sample_lists/sample_table_unmerged.tsv -n $K | tail -n 1 | cut -f 4`
  LANE=`head /workdir/batch-effect/sample_lists/sample_table_unmerged.tsv -n $K | tail -n 1 | cut -f 2`
  SEQID=`head /workdir/batch-effect/sample_lists/sample_table_unmerged.tsv -n $K | tail -n 1 | cut -f 3`
  DATATYPE=`head /workdir/batch-effect/sample_lists/sample_table_unmerged.tsv -n $K | tail -n 1 | cut -f 6`
  
  PREFIX=$SAMPLE'_'$SEQID'_'$LANE'_'$DATATYPE'_bt2_gadMor3'
  
  if [ $DATATYPE = se ]; then
    INPUTPATH='/workdir/cod/greenland-cod/bam/'
  else
    INPUTPATH='/workdir/batch-effect/bam/'
  fi
  #echo $INPUTPATH$PREFIX'.bam'
  #echo $OUTPUTPATH$PREFIX'_sorted.bam'
  samtools sort $INPUTPATH$PREFIX'.bam' -o $OUTPUTPATH$PREFIX'_sorted.bam' -@ 24
done

```

```{bash eval=F}
nohup bash /workdir/batch-effect/scripts/sort_raw_bam.sh > /workdir/batch-effect/nohups/sort_raw_bam.nohup &
```

#### Merge four samples that were sequenced in multiple lanes

```{bash eval=F}
## Merge four samples that were sequenced in multiple lanes
samtools merge /workdir/batch-effect/bam/QQL2011_842_merged_merged_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_842_14247X225_1_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_842_14247X47_2_se_bt2_gadMor3_sorted.bam -@ 24
samtools merge /workdir/batch-effect/bam/QQL2011_844_merged_merged_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_844_14247X226_1_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_844_14247X49_2_se_bt2_gadMor3_sorted.bam -@ 24
samtools merge /workdir/batch-effect/bam/QQL2011_846_merged_merged_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_846_14247X232_1_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_846_14247X61_2_se_bt2_gadMor3_sorted.bam -@ 24
samtools merge /workdir/batch-effect/bam/QQL2011_852_merged_merged_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_852_14247X228_1_se_bt2_gadMor3_sorted.bam /workdir/batch-effect/bam/QQL2011_852_14247X40_2_se_bt2_gadMor3_sorted.bam -@ 24
```

## Deduplicate

I slightly modified the original script (`/workdir/data-processing/scripts/deduplicate_clipoverlap.sh`) so that the `minq20` part is no longer in the file name.

```{bash eval=F}
## SE
nohup bash /workdir/batch-effect/scripts/deduplicate_clipoverlap.sh \
/workdir/batch-effect/sample_lists/bam_list_merged_se.txt \
/workdir/batch-effect/sample_lists/sample_table_merged.tsv \
/workdir/batch-effect/ \
gadMor3 \
> /workdir/batch-effect/nohups/deduplicate_se.nohup &
## PE
nohup bash /workdir/batch-effect/scripts/deduplicate_clipoverlap.sh \
/workdir/batch-effect/sample_lists/bam_list_merged_pe.txt \
/workdir/batch-effect/sample_lists/sample_table_merged.tsv \
/workdir/batch-effect/ \
gadMor3 \
> /workdir/batch-effect/nohups/deduplicate_clipoverlap_pe.nohup &
```

## Indel realignment with PE and SE data combined

```{bash eval=F}
cp /workdir/batch-effect/sample_lists/bam_list_dedup_overlapclipped.txt \
/workdir/batch-effect/sample_lists/bam_list_dedup_overlapclipped.list

nohup bash /workdir/data-processing/scripts/realign_indels.sh \
/workdir/batch-effect/sample_lists/bam_list_dedup_overlapclipped.list \
/workdir/batch-effect/ \
/workdir/cod/reference_seqs/gadMor3.fasta \
gadMor3 \
> /workdir/batch-effect/nohups/realign_indels.nohup &
```

## Get genotype likelihoods

#### polyG trimmed PE samples (original)

This is with PE samples that went through polyG trimming instead of sliding window trimming.

```{bash eval=F}
## original setup
cd /workdir/cod/greenland-cod/
/workdir/programs/angsd0.931/angsd/angsd \
-b sample_lists/bam_list_realigned_mincov_contamination_filtered_batch_effect.txt \
-anc /workdir/cod/reference_seqs/gadMor3.fasta \
-out angsd/bam_list_realigned_mincov_contamination_filtered_batch_effect_mindp2_maxdp661_minind2_minq20 \
-GL 1 -doGlf 2 -doMaf 1 -doMajorMinor 3 -doCounts 1 -doDepth 1 -dumpCounts 1 \
-P 32 -setMinDepth 2 -setMaxDepth 661 -minInd 2 -minQ 20 -minMaf 0.01 \
-sites angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.txt \
-rf angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.chrs \
>& nohups/get_gl_bam_list_realigned_mincov_contamination_filtered_batch_effect.log &

## more stringent filtering
cd /workdir/cod/greenland-cod/
/workdir/programs/angsd0.931/angsd/angsd \
-b sample_lists/bam_list_realigned_mincov_contamination_filtered_batch_effect.txt \
-anc /workdir/cod/reference_seqs/gadMor3.fasta \
-out angsd/bam_list_realigned_mincov_contamination_filtered_batch_effect_mindp2_maxdp661_minind2_stringent_filter \
-GL 1 -doGlf 2 -doMaf 1 -doMajorMinor 3 -doCounts 1 -doDepth 1 -dumpCounts 1 \
-P 32 -setMinDepth 2 -setMaxDepth 661 -minInd 2 -minMaf 0.01 \
-minQ 33 -minMapQ 25 -uniqueOnly 1 \
-sites angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.txt \
-rf angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.chrs \
>& nohups/get_gl_bam_list_realigned_mincov_contamination_filtered_batch_effect_stringent_filter.log &
```

#### Sliding window trimmed PE samples (new)

This is with PE samples that went through sliding window trimming.

I will first try the same LD-trimmed SNP list from the Greenland cod project

```{bash eval=F}
cd /workdir/batch-effect/
## Run ANGSD as usual (with a minMapQ filter of 20)
nohup /workdir/programs/angsd0.931/angsd/angsd \
-b sample_lists/bam_list_realigned.txt \
-anc /workdir/cod/reference_seqs/gadMor3.fasta \
-out angsd/bam_list_realigned \
-GL 1 -doGlf 2 -doMaf 1 -doMajorMinor 3 -doCounts 1 -doDepth 1 -dumpCounts 1 \
-P 16 -setMinDepth 2 -setMaxDepth 661 -minInd 2 -minQ 20 -minMapQ 20 -minMaf 0.01 \
-doIBS 2 -makematrix 1 -doCov 1 \
-sites /workdir/cod/greenland-cod/angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.txt \
-rf /workdir/cod/greenland-cod/angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.chrs \
>& nohups/get_gl_bam_list_realigned.log &
## Get the matrices with more stringent filtering
nohup /workdir/programs/angsd0.931/angsd/angsd \
-b sample_lists/bam_list_realigned.txt \
-anc /workdir/cod/reference_seqs/gadMor3.fasta \
-out angsd/bam_list_realigned_stringent \
-doMajorMinor 3 -doCounts 1 -doDepth 1 -dumpCounts 1 \
-P 16 -setMinDepth 40 -setMaxDepth 326 -minInd 20 -minQ 33 -minMapQ 30 \
-doIBS 2 -makematrix 1 -doCov 1 \
-sites /workdir/cod/greenland-cod/angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.txt \
-rf /workdir/cod/greenland-cod/angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.chrs \
>& nohups/get_matrix_stringent_bam_list_realigned.log &
## Get depth count without mapping quality filter
nohup /workdir/programs/angsd0.931/angsd/angsd \
-b sample_lists/bam_list_realigned.txt \
-anc /workdir/cod/reference_seqs/gadMor3.fasta \
-out angsd/bam_list_realigned_anymapq \
-doCounts 1 -doDepth 1 -dumpCounts 1 \
-P 16 -setMinDepth 2 -minInd 2 -minQ 20 \
-sites /workdir/cod/greenland-cod/angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.txt \
-rf /workdir/cod/greenland-cod/angsd/global_snp_list_bam_list_realigned_mincov_contamination_filtered_mindp151_maxdp661_minind102_minq20_downsampled_unlinked.chrs \
>& nohups/get_depth_anymapq_bam_list_realigned.log &
```

## Run PCAngsd

#### Sliding window trimmed PE samples (new)

```{bash eval=F}
nohup python2 /workdir/programs/pcangsd/pcangsd.py \
-beagle /workdir/batch-effect/angsd/bam_list_realigned.beagle.gz \
-minMaf 0.05 \
-threads 8 \
-o /workdir/batch-effect/angsd/bam_list_realigned_pcangsd \
> /workdir/batch-effect/nohups/run_pcangsd.nohup &
```
